HUME: Human UCCA-based Machine-Translation Evaluation

### 2016-03-14 paper/main.tex

  Recent interest in semantics-aware MT systems underscores te importance of
  semantic evaluation for Machine Translation (MT) systems. 
  A semantic measure, sensitive to possible semantic discrepencies
  between the translation output and the source, is likely support the construction
  and tuning of structure-aware MT, much in the same way that string-based measures,
  such as BLEU, support advances with shallower MT systems.  
  We present a novel human semantic evaluation measure for MT, Human
  UCCA-based MT evaluation (HUME), building on the UCCA semantic representation scheme.
  Making progress over previous work, the proposed measure takes into account
  a wider range of semantic phenomena and does not rely on semantic annotation
  of the MT output itself, whose interpretation is often unclear.
  We experiment with four language pairs, demonstrating the broad applicability of
  the measure, as well as its reliability, reflected in the inter-annotator agreement
  scores.

@@

Human judgement is the central, and perhaps most fundamental, criterion for
estimating the quality of an MT system.
Nevertheless, common measures for MT evaluation, such as adequacy and frequency judgements
or the relative ranking of several possible translations, are problematic in two ways.
First, as the quality of translation is determined by multiple factors, it is difficult
to assign a single number that quantifies the quality of the whole entire sentence. This
is indeed reflected in the diminishing inter-annotator agreement rates of human ranking measures
as a function of the sentence length \cite{Bojar:2011}.
Second, a sentence-level quality score does not indicate what parts of the sentence
are erroneously translated, and hence cannot inform developers in overcoming these errors.

These problems were partially addressed by measures that decompose over sub-parts of the sentence,
most commonly over its words or n-grams, quantifying the overlap of the sub-parts with a reference
translation. More recent work on MT evaluation proposed measures
that decompose over semantically defined units,
quantifying the similarity of the output and the reference in terms of
their verb argument structure (the most notable of these measures is
MEANT \cite{lo2011structured}).
While allowing to localize semantic errors in the translation,
these measures' focus on verb argument structures misses out on many pervasive semantic phenomena,
such as inter-clause linkage and nominal argument structures, and thus only provides
a limited perspective on the semantic similarity between the output and reference.
Moreover, assigning semantic structure to MT output, especially when low in quality,
is difficult and arguably ill-defined.

In this work we propose the Human UCCA-based MT Evaluation ({\it HUME}) metric,
a human evaluation measure that decomposes over the semantic units of the sentence.
Semantic units are defined according to the 
UCCA scheme \cite{abend2013universal}, an appealing candidate for semantic analysis,
due to its cross-linguistic applicability, support for rapid annotation, and coverage
of many fundamental semantic phenomena, such as verbal, nominal and adjectival
argument structures and the inter-relations between them.

HUME operates by aggregating human assessments of the translation quality of individual
semantic units in the source sentence.
Importantly, HUME only requires semantically annotating the source sentence,
thus avoiding the annotatation of machine-generated text 
and allowing re-use of the source semantic annotation for measuring the quality
of different translations of the same source sentence.
By basing the measure on a direct comparison
of the source and translation we avoid the difficulties engendered by using
reference translations, albeit at the cost of imposing the additional requirement
that the evaluator be proficient in both the source and target languages.
See \secref{sec:hmeant_comp} for further discussion.

We conduct experiments with four language pairs, English-German, English-Polish,
English-Romanian and English-Czech, and report efficiency (time of annotation) and
inter-annotator agreement scores, as well as qualitative comparison of the resulting
HUME scores with crowd-sourced adequacy judgements. Results show that HUME obtains 
agreements which are on par with those of standard sentence-level ranking measures,
but unlike them, HUME's agreement does not diminish with the sentence length.
The paper concludes by outlining ongoing work to construct semi- and fully-automatic variants of HUME.

@@

\paragraph{Machine Translation Evaluation.}
%MT evaluation has been traditionally pursued through human-based
%evaluation that quantifies the adequacy of the translation in preserving the meaning 
%of the source sentence, and its fluency as a sentence in the target language \cite{lopez2008statistical}.
Human evaluation is the standard measure, and is
estimated either in terms of the relative ranking of the outputs of multiple systems
e.g., in the WMT shared tasks \cite{bojar2015findings}, or by assigning
indiviaul adequacy/fluency scores to each sentence, a procedure recently improved
by \newcite{graham2015improving}.
However, while providing the gold standard for MT evaluation, human evaluation is not a scalable solution,
and is unfeasible for a wide variety of systems or hyper-parameter settings are to be evaluated.
which has motivated the construction of cheaper methods that correlate with human
judgement.

Scalability is addressed by scores that decompose over the sub-parts of
the translation, and quantify how many of these sub-parts appear in a reference translation,
whose validity has been manually confirmed. The most commonly used measures decompose
over n-grams or even individual words, e.g., 
the BLEU \cite{Papineni:2002}, NIST \cite{Doddington:2002} and METEOR measures \cite{Banerjee:2005}.
Other methods employ a string edit distance for determining the similarity between the
reference and translation \cite{snover2006study}.
While these measures stimulated much progress in MT research by allowing
to evaluate experiments no a massive scale, they 
decompose over the words and n-grams of the sentence,
and are thus better suited for localizing errors
in the output of shallow MT methods, and are less telling for identifying the semantic
nature of translation errors in the output.

@@

To address this shortcoming, more more recent work quantifies
the similarity of the reference and translation output in terms
of their {\it semantic structure}. \newcite{liu2005syntactic} took a syntactic approach, 
quantifying the similarity between the syntactic structures of the reference and translation.
\newcite{owczarzak2007evaluating} took a similar approach using lexical-functional grammar structures.
\newcite{gimenez2007linguistic} proposed an evaluation measure that incorapotes features
that capture the overlap between the output and reference in terms of their
semantic (argument structure phenomena), lexical and morphosyntactic features.

Perhaps the most notable attempt at semantic MT evaluation is MEANT and
its human variant HMEANT, which quantifies the similarity between
the reference and translation in terms of the overlap in
their verbal argument structures structures, and associated semantic roles.
HMEANT shares much of HUME's motivation for providing a
measure that decomposes over semantic units, rather than strings, but focuses on verbal
argument structures, ignoring other pervasive phenomena such as nominal predicates, inter-clause
relations and copula clauses. HUME further diverges from both MEANT and the more shallow BLEU and
METEOR measures, in not requiring a reference. Instead,
it compares the source directly with the output
translation. We see this difference as crucial where semantic analysis is concerned, as it
is often impossible to assign a semantic structure to a low quality MT output sentence.
See \secref{sec:hmeant_comp} for a more elaborate discussion of the differences between HUME
and HMEANT.

@@

% UCCA Motivation
UCCA is a natural candidate for defining a semantic
MT evaluation measure for a number of reasons. 
First, it defines a coarse-grained and intuitive set of distinctions definitions,
and consequently can be reliably annotated by non-experts, after as little as two hours
of training \cite{marinotti2014}.
Second, UCCA is a cross-linguistically applicable scheme, seeking to represent what is shared between languages,
by building on linguistic typological theory, and specifically on ``Basic Linguistic Theory''
\cite{Dixon:10a,Dixon:10b,Dixon:12}, one of the most widely used frameworks for linguistic description.
Indeed, UCCA's cross-linguistic applicability has been demonstrated through its
application to annotation in English, French, German and Czech\oa{Did we do experiments on Czech?}. 
Third, the scheme has been shown to be semantically stable
in translation: UCCA annotations of translated text usually contain the same set of relationships
\cite{sulem2015conceptual}. This is an indication that the scheme reflects
a common layer of representation that in a correct translation
is mostly shared between the translation and the source. 
We note that that verbal argument structures are less stable in this sense, as a verbal
argument structure may be translated or paraphrased to a nominal or adjectival
argument structure, and are thus less motivated as a basis for a semantic measure.
For example, ``after he graduated, Tom moved to NYC''
conveys a similar meaning to ``after graduation, Tom moved to NYC''.

While most broad-coverage semantic work focused on shallow semantic structures,
predominantly argument structures annotated with their semantic roles, much recent
interest has been devoted to defining more complete semantic representation schemes.
The Abstract Meaning Representation (AMR) scheme \cite{banarescu2013abstract}
shares much of UCCA's motivation for defining a more complete semantic annotation.
However, using AMR is not optimal for defining a decomposition of a sentence into semantic
units as it does not ground its semantic symbols in the text,
thus not straightwardly allowing a decomposition of the sentence into sub-units.
Moreover, AMR is more fine-grained than UCCA and consequently more difficult to annotate.

@@

We do not distinguish fluency and adequacy judgements, as they are correlated anyway \cite{ccb}.

@@

There are two manual steps in the annotation process. The first is to create UCCA annotations for the source sentence,
following the UCCA guidelines. Since this is is fully explained elsewhere \cite{abend2013universal}, we will not go into detail
here, except to say that in our case the source sentences were all in English, and  the annotation was performed by four
computational linguists. After we have the UCCA annotation on 
the source, we use the word-to-word alignments from the MT system in order to project this annotation to the MT output, and
we used native speakers of the target language in order to annotate this output. Since these annotators have to
judge the quality of the translation, they also needed to be fluent in English.\bh{Somewhere we could explain that we
could do a similar annotation with monolinguals, although it entails aligning reference and MT output}

@@

@@

We conducted an analysis of the English UCCA Wikipedia
corpus\footnote{\url{http://www.cs.huji.ac.il/~oabend/ucca.html}}, comprising
of 5324 sentences,
in order to assess the frequency of some of the major phenomena HMEANT fails to address,
Coppula cluases are treated in HMEANT simply as instances of the main verb ``be'', which
generally does not convey the meaning of these clauses. They appear in 21.7\% of the sentences,
according to a conservative estimates that only considers non-auxilliary instances of ``be''.
Nominative argument structures, which are completely ignored under HMEANT are in fact highly
pervasive, appearing in 48.7\% of the sentences. Linkers expressing an inter-relation between
clauses (mainly discourse markers and coordinating conjunctions) appear in 56\% of the
sentences\footnote{Argument structures and linkers are explicitly marked in UCCA. We identified
  non-auxilliary instances of ``be'' and nouns using the NLTK standard tagger.}.
We are not aware of any empirical evaluation that demonstrates that verb argument structure,
taken alone, capture the crux of the sentence semantics meanings, and so the coverage of
further phenomena by HUME will first allow to determine empirically
(rather than stipulate), what semantic structures are important
to preserve for MT translation, and which are less important

>>>>

  Recent interest in semantics-aware MT systems underscores the importance of
  semantic evaluation for Machine Translation (MT) systems. 
  A semantic measure, sensitive to possible semantic discrepencies
  between the translation output and the source, is likely support the construction
  and tuning of structure-aware MT, much in the same way that string-based measures,
  such as BLEU, enabled progress in shallower MT.
  We present a novel human semantic evaluation measure for MT, Human
  UCCA-based MT evaluation (HUME), building on the UCCA semantic representation scheme.
  Making progress over previous work, the proposed measure takes into account
  a wider range of semantic phenomena and does not rely on semantic annotation
  of the MT output itself, whose interpretation is often unclear.
  We experiment with four language pairs, demonstrating HUME's broad applicability,
  as well as its reliability, reflected in its inter-annotator agreement rates.

@@

Human judgement is the central, and perhaps most fundamental, criterion for
estimating the quality of an MT system.
Nevertheless, common measures for MT evaluation, such as adequacy and fluency judgements
or the relative ranking of several possible translations, are problematic in two ways.
First, as the quality of translation is determined by multiple factors, it is difficult
to assign a single number that quantifies the quality of the entire sentence. This
is indeed reflected in the diminishing inter-annotator agreement rates of human ranking measures
as a function of the sentence length \cite{Bojar:2011}.
Second, a sentence-level quality score does not indicate what parts of the sentence
are erroneously translated, and hence cannot inform developers in overcoming these errors.

These problems were partially addressed by measures that decompose over sub-parts of the evaluated
translation sentence (henceforth, {\it translation}),
most commonly over its words or n-grams,
quantifying the overlap of its sub-parts relative a reference translation.
More recent work on MT evaluation proposed measures
that decompose over semantically defined units,
quantifying the similarity of the output and the reference in terms of
their verb argument structure (the most notable of these measures is
MEANT \cite{lo2011structured}).
While allowing to localize semantic errors in the translation,
these measures' focus on verb argument structures misses out on many pervasive semantic phenomena,
such as inter-clause linkage and nominal argument structures, and thus only provides
a limited perspective on the semantic similarity between the output and reference.
Moreover, assigning semantic structure to MT output, especially when low in quality,
is difficult and arguably ill-defined.

We propose the Human UCCA-based MT Evaluation ({\it HUME}) metric,
a human evaluation measure that decomposes over the semantic units of the sentence.
Semantic units are defined according to the 
UCCA scheme \cite{abend2013universal}, an appealing candidate for semantic analysis,
due to its cross-linguistic applicability, support for rapid annotation, and coverage
of many fundamental semantic phenomena, such as verbal, nominal and adjectival
argument structures and the inter-relations between them.

HUME operates by aggregating human assessments of the translation quality of individual
semantic units in the source sentence.
Importantly, HUME only requires semantically annotating the source sentence,
thus avoiding the annotatation of machine-generated text 
and allowing re-use of the source semantic annotation for measuring the quality
of different translations of the same source sentence.
By basing the measure on a direct comparison
of the source and translation we avoid the difficulties engendered by using
reference translations, albeit at the cost of imposing the additional requirement
that the evaluator be proficient in both the source and target languages.
See \secref{sec:hmeant_comp}.

We conduct experiments with four language pairs: English-German, English-Polish,
English-Romanian and English-Czech, and report efficiency (time of annotation) and
inter-annotator agreement scores, as well as qualitative comparison of the resulting
HUME scores with crowd-sourced adequacy judgements. Results show that HUME obtains 
agreement scores which are on par with those of standard sentence-level ranking measures,
but unlike them, HUME's scores do not decrease with logner sentences.
The paper concludes by outlining ongoing work to construct semi- and
fully-automatic variants of HUME.

@@

\paragraph{Machine Translation Evaluation.}
%MT evaluation has been traditionally pursued through human-based
%evaluation that quantifies the adequacy of the translation in preserving the meaning 
%of the source sentence, and its fluency as a sentence in the target language \cite{lopez2008statistical}.
Human evaluation is generally determined
either according to the relative ranking of the outputs of multiple systems
e.g., in the WMT shared tasks \cite{bojar2015findings}, or by assigning
indiviaul adequacy/fluency scores to each translation, a procedure recently improved
by \newcite{graham2015improving}.
However, while providing the gold standard for MT evaluation, human evaluation is not a scalable solution,
and is unfeasible where a wide variety of systems or parameter settings are to be evaluated.
% has motivated the construction of cheaper methods that correlate with human
%judgement.

Scalability is addressed by employing automatic and semi-automatic approximations of human
judgements. Most commonly, such scores decompose over the sub-parts of
the translation, and quantify how many of these sub-parts appear in a reference translation,
whose validity has been manually confirmed. This decomposition further allows system developers
to localize the errors within the translation.
The most commonly used measures decompose over n-grams or even individual words, e.g., 
the BLEU \cite{Papineni:2002}, NIST \cite{Doddington:2002} and METEOR measures \cite{Banerjee:2005}.
Another common approach is to determine the similarity between the reference and translation
by applying string edit distance \cite{snover2006study}.
While these measures stimulated much progress in MT research by allowing
the evaluation of massive-scale experiments,
their decomposition over words and n-grams is better suited for localizing errors
in the output of shallow MT methods, and is less instructive in identifying the semantic
nature of the errors.

@@

In order to address this shortcoming, more recent work quantified
the similarity of the reference and translation in terms
of their {\it semantic structure}. \newcite{liu2005syntactic} took a syntactic approach, 
quantifying the similarity between the syntactic structures of the reference and translation.
\newcite{owczarzak2007evaluating} took a similar approach using lexical-functional grammar structures.
\newcite{gimenez2007linguistic} proposed to aggregate multiple types of information,
capturing the overlap between the output and reference in terms of their
semantic (predicate-argument structures), lexical and morphosyntactic features.

Perhaps the most notable attempt at semantic MT evaluation is MEANT and
its human variant HMEANT, which quantifies the similarity between
the reference and translation in terms of the overlap in
their verbal argument structures structures, and associated semantic roles.
We discuss the differences between HMEANT and HUME in \secref{sec:hmeant_comp}.

@@

% UCCA Motivation
UCCA is a natural candidate for defining a semantic
MT evaluation measure for a number of reasons. 
First, it defines a coarse-grained and intuitive set of categories,
and consequently can be reliably annotated by non-experts, after as little as two hours
of training \cite{marinotti2014}.
Second, UCCA is a cross-linguistically applicable scheme, seeking to represent what is shared between languages,
by building on linguistic typological theory, and specifically on ``Basic Linguistic Theory''
\cite{Dixon:10a,Dixon:10b,Dixon:12}, one of the most widely used frameworks for linguistic description.
Indeed, UCCA's cross-linguistic applicability has been demonstrated through its
application to annotation in English, French, German and Czech\oa{Did we do experiments on Czech?}. 
Third, the scheme has been shown to be semantically stable
in translation: UCCA annotations of translated text usually contain the same set of relationships
\cite{sulem2015conceptual}. This is an indication that the scheme reflects
a common layer of representation that in a correct translation
is mostly shared between the translation and the source. 
We note that that verbal argument structures are less stable in this sense, as a verbal
argument structure may be translated or paraphrased to a nominal or adjectival
argument structure, and are thus less suitable as a basis for a semantic measure.
For example, ``after he graduated, Tom moved to NYC''
conveys a similar meaning to ``after graduation, Tom moved to NYC''.

While most broad-coverage semantic work focused on shallow semantic structures,
predominantly argument structures annotated with their semantic roles, much recent
interest has been devoted to defining more complete semantic representation schemes.
The Abstract Meaning Representation (AMR) scheme \cite{banarescu2013abstract}
shares much of UCCA's motivation for defining a more complete semantic annotation.
However, using AMR is not optimal for defining a decomposition of a sentence into semantic
units as it does not ground its semantic symbols in the text,
thus not straightwardly allowing a decomposition of the sentence into sub-units.
Moreover, AMR is more fine-grained than UCCA and consequently more difficult to annotate.\oa{should we
  add that it is less cross-linguistically stable? maybe that's an overkill..}

@@

%We do not distinguish fluency and adequacy judgements, as they are correlated anyway \cite{ccb}.

@@

There are two manual steps in the annotation process. The first is to create UCCA annotations for the source sentence,
following the
UCCA guidelines\footnote{All UCCA-related resources can be found here: \url{http://www.cs.huji.ac.il/~oabend/ucca.html}}. Source sentences are all in English, and the annotation was performed by four
computational linguists\oa{were there no other annotators from Prague?}. After we have the UCCA annotation on 
the source, we use the word-to-word alignments from the MT system in order to project this annotation to the MT output,
and use native speakers of the target language in order to annotate this output. Since these annotators have to
judge the quality of the translation, they also needed to be fluent in English.\bh{Somewhere we could explain that we
could do a similar annotation with monolinguals, although it entails aligning reference and MT output}

@@

HMEANT shares much of HUME's motivation for providing a
measure that decomposes over semantic units, rather than strings, but focuses on verbal
argument structures, ignoring other pervasive phenomena such as nominal predicates, inter-clause
relations and copula clauses.
HUME further diverges from both MEANT and the shallower BLEU and
METEOR measures, in not requiring a reference.
Instead, it compares the source directly with the output
translation. We see this difference as crucial where semantic analysis is concerned, as it
is often impossible to assign a semantic structure to a low quality MT output sentence.
See \secref{sec:hmeant_comp} for a more elaborate discussion of the differences between HUME
and HMEANT.

@@

We conducted an analysis of the English UCCA Wikipedia
corpus, comprising
of 5324 sentences,
in order to assess the frequency of some of the major phenomena HMEANT fails to address,
Coppula cluases are treated in HMEANT simply as instances of the main verb ``be'', which
generally does not convey the meaning of these clauses. They appear in 21.7\% of the sentences,
according to a conservative estimates that only considers non-auxilliary instances of ``be''.
Nominative argument structures, which are completely ignored under HMEANT are in fact highly
pervasive, appearing in 48.7\% of the sentences. Linkers expressing an inter-relation between
clauses (mainly discourse markers and coordinating conjunctions) appear in 56\% of the
sentences\footnote{Argument structures and linkers are explicitly marked in UCCA. We identified
  non-auxilliary instances of ``be'' and nouns using the NLTK standard tagger.}.
We are not aware of any empirical evaluation that demonstrates that verb argument structure,
taken alone, capture the crux of the sentence semantics meanings, and so the coverage of
further phenomena by HUME will first allow to determine empirically
(rather than stipulate), what semantic structures are important
to preserve for MT translation, and which are less important
